The Evolution of Phenotypically Invariant Genetic Networks

### 2021-05-06 kalman_walk_paper.tex
.

Now consider the perturbed systems $(A(\epsilon), B(\epsilon), C(\epsilon)) = (A(0), B(0), C(0)) + \epsilon (U, V, W)$,
and let $\h_\epsilon(t) = C(\epsilon) e^{A(\epsilon) t} B(\epsilon)$.
If two populations with systems $(A(0), B(0), C(0))$ and $(A(\epsilon), B(\epsilon), C(\epsilon))$ interbreed,
then the mean resulting system is $(A({\epsilon/2}), B({\epsilon/2}), C({\epsilon/2}))$,
but with greater variance in the $F_2$ than the $F_1$ offspring (see Figure \ref{fig:conceptual_fig}).
Indeed, if the difference between (say) $A(0)_{ij}$ and $A(\epsilon)_{ij}$
is due to $n$ fixed differences with additive effect size $\epsilon/(2n)$ each,
then the standard deviation of $A_{ij}$ among the $F_1$ offspring is equal to that
within each parent population, while among the $F_2$ offspring it is $\epsilon / \sqrt{8 n}$.
This implies that the impulse response, $h(t)$, of an $F_2$ will also differ from optimum
by something of order $\epsilon$.
However, since the displacement between the populations lies along a ridge,
the $F_1$ offspring are closer:
the Taylor expansions for $h$ about both 0 and $\epsilon$ are
\begin{align}
    h_{\epsilon/2}(t)
    &= h_0(t)
        + \frac{\epsilon}{2} \partial_\epsilon h_0(t)
        + \frac{\epsilon^2}{8} \partial_\epsilon^2 h_0(t)
        + O(\epsilon^3) \\
    \text{and} \quad
    &= h_\epsilon(t)
        - \frac{\epsilon}{2} \partial_\epsilon h_\epsilon(t)
        + \frac{\epsilon^2}{8} \partial_\epsilon^2 h_\epsilon(t)
        + O(\epsilon^3)
\end{align}
Since $h_\epsilon(t) = h_0(t)$ by assumption,
and $\partial_\epsilon h_\epsilon(t) = \partial_\epsilon h_0(t) + \epsilon \partial_\epsilon^2 h_0(t) + O(\epsilon^2)$,
combining these expressions we get that
\begin{align}
    h_{\epsilon/2}(t)
    &= h_0(t)
        - \frac{\epsilon^2}{8} \partial_\epsilon^2 h_0(t)
        + O(\epsilon^3) .
\end{align}
In other words, if the parental populations are both optimal but differ by $\epsilon$,
then the $F_1$ offspring differ from optimal by something of order $\epsilon^2$,
but the $F_2$ by something of order $\epsilon$.

>>>>

Now consider the perturbed systems $(A(\epsilon), B(\epsilon), C(\epsilon)) = (A(0), B(0), C(0)) + \epsilon (U, V, W)$,
and let $\h_\epsilon(t) = C(\epsilon) e^{A(\epsilon) t} B(\epsilon)$.
If two populations with systems $(A(0), B(0), C(0))$ and $(A(\epsilon), B(\epsilon), C(\epsilon))$ interbreed,
then the mean resulting system is $(A({\epsilon/2}), B({\epsilon/2}), C({\epsilon/2}))$,
but with greater variance in the $F_2$ than the $F_1$ offspring (see Figure \ref{fig:conceptual_fig}).
Indeed, if the difference between (say) $A(0)_{ij}$ and $A(\epsilon)_{ij}$
is due to $n$ fixed differences with absolute additive effect size $\epsilon/\sqrt{n}$ each,
then the standard deviation of $A_{ij}$ among the $F_1$ offspring is equal to that
within each parent population, while among the $F_2$ offspring it is $\epsilon/2$.
(This many differences of this size would be expected if the difference was due to drift, for instance.)
This implies that the impulse response, $h(t)$, of an $F_2$ will also differ from optimum
by something of order $\epsilon$.
However, since the displacement between the populations lies along a ridge,
the $F_1$ offspring are closer:
the Taylor expansions for $h$ about both 0 and $\epsilon$ are
\begin{align}
    h_{\epsilon/2}(t)
    &= h_0(t)
        + \frac{\epsilon}{2} \partial_\epsilon h_0(t)
        + \frac{\epsilon^2}{8} \partial_\epsilon^2 h_0(t)
        + O(\epsilon^3) \\
    \text{and} \quad
    &= h_\epsilon(t)
        - \frac{\epsilon}{2} \partial_\epsilon h_\epsilon(t)
        + \frac{\epsilon^2}{8} \partial_\epsilon^2 h_\epsilon(t)
        + O(\epsilon^3)
\end{align}
Since $h_\epsilon(t) = h_0(t)$ by assumption,
and $\partial_\epsilon h_\epsilon(t) = \partial_\epsilon h_0(t) + \epsilon \partial_\epsilon^2 h_0(t) + O(\epsilon^2)$,
combining these expressions we get that
\begin{align}
    h_{\epsilon/2}(t)
    &= h_0(t)
        - \frac{\epsilon^2}{8} \partial_\epsilon^2 h_0(t)
        + O(\epsilon^3) .
\end{align}
In other words, if the parental populations are both optimal but differ by $\epsilon$,
then the $F_1$ offspring differ from optimal by something of order $\epsilon^2$,
but the $F_2$ by something of order $\epsilon$.
